\sysfull

### 2018-08-04 paper/asplos19/background.tex
revise intro

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
%
Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
Figure~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation.
If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).

>>>>

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
%
Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
Fig.~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation.
If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).

### 2018-08-04 paper/asplos19/intro.tex
revise intro

\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a

@@

statically decompose a program into a collection of {\em tasks}.  A task can

@@

Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.

@@

A programmer may thus create very large tasks in an attempt to reduce
overhead by minimizing the number of task transitions. However, a very large
task may require more energy to complete than a device's fixed hardware energy
buffer can hold which may lead to a program non-termination problem. To eliminate this risk,
existing systems require the programmer to decompose a program
into small tasks, all of which complete using only buffered energy. These
constraints on task sizing leave an unsatisfying dilemma: large, efficient
tasks that risk non-termination, or small tasks that are guaranteed to
complete, but incur a high task transition and commit overhead. In this work,
we propose a third way, using a novel technique called {\em adaptive execution by dynamic task
coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.

@@

This paper introduces a new system called \sys, which executes a task-based

@@

\sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
with no commit overhead between them, instead performing
task-end commit actions at the end of the second task only.
When there is sufficient energy to execute both tasks, two distinct, small
tasks are effectively executed as a single large task, preserving the atomicity
of both. If power fails during a coalesced task, execution restarts from the
{\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution.
On the other hand, a large task that requires more energy to complete than a device buffered energy
is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
\textbf{Challenges and Contributions} 
\begin{itemize}
\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly? \textit{contribution---1} \sys addresses this challenge by using its \textit{recent execution history} as a metric to estimate the environment energy conditions. Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully. 
\item\textit{challenge} 2) Given dynamic variables dependencies, how to ensure efficient data protection against power interrupts? \textit{contribution---2} \sys ensures the consistency of the data by using a novel
approach called \emph{dynamic group privatization}. Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e. protecting all global variables) in each dynamic task transition. \sys, therefore, utilizes a real-time data tracking to enable partial data protection on a task transition. Individual variables tacking, however, slows down a system dramatically. Therefore, \sys  relies on a coarse locality principle and protects the global variables in a batch. Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
\item\textit{challenge} 3) A static task decomposition model assumes that each
single task can execute to completion. If the hardware energy buffer provides
inadequate energy to execute each single task to completion, a program will not
terminate~\cite{cleancut_2018}. under such conditions how to enable the dynamic execution model to progress on a sub-task level? \textit{contribution---3} To avoid non-termination under adversarial
energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
Partial execution avoids non-termination by committing the intermediate state of a long-running
task that has repeatedly failed and restarted. Partial execution violates
task atomicity, but preserves forward progress; if a programmer knows that
task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
\end{itemize}
Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} provides an overview of \sys, while
Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism. Dynamic group privatization mechanism is provided in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's methodology and evaluation, respectively. Section~\ref{sec:related_work} puts \sys in the
context of related work and Section~\ref{sec:conclusions} concludes and discusses
future work.

>>>>

\textbf{Task Decomposition of Intermittent Programs.} Task-based programming and execution models require a

@@

statically decompose a program into a collection of tasks.  A \emph{task}, a top level function, can

@@

Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and
shows how tasks transitions can impose a run-time overhead. At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.

@@

A programmer may thus create very large tasks in an attempt to reduce task transitions
overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy
buffer can hold which may lead to a task non-termination problem. To eliminate this risk,
existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress. These constraints on task sizing leave an unsatisfying dilemma: large, efficient tasks that risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead. In this work,
we propose a third way, using a novel technique called \emph{adaptive task size execution by task coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.

@@

This paper introduces a new system called \emph{\sys}, which executes a task-based

@@

\sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce} (top)) or splits them (Fig.~\ref{fig:coalesce} (bottom)) based on the estimated energy availability. 
\sys does not requires hardware support for energy estimation to make its scheduling decisions.

% \sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
% with no commit overhead between them, instead performing task-end commit actions at the end of the second task only.
% %
% When there is sufficient energy to execute both tasks, two distinct, small
% tasks are effectively executed as a single large task, preserving the atomicity
% of both. If power fails during a coalesced task, execution restarts from the
% {\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
% illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution.
% On the other hand, a large task that requires more energy to complete than a device buffered energy
% is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).

\textbf{Challenges and Contributions} 
\begin{itemize}
\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly? \textit{contribution---1} \sys addresses this challenge by using its \textit{recent execution history} as a metric to estimate the environment energy conditions. Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully. 
%
\item\textit{challenge} 2) Given dynamic variables dependencies, how to ensure efficient data protection against power interrupts? \textit{contribution---2} \sys ensures the consistency of the data by using a novel
approach called \emph{dynamic group privatization}. Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e. protecting all global variables) in each dynamic task transition. \sys, therefore, utilizes a real-time data tracking to enable partial data protection on a task transition. Individual variables tacking, however, slows down a system dramatically. Therefore, \sys  relies on a coarse locality principle and protects the global variables in a batch. Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
%
\item\textit{challenge} 3) A static task decomposition model assumes that each
single task can execute to completion. If the hardware energy buffer provides
inadequate energy to execute each single task to completion, a program will not
terminate~\cite{cleancut_2018}. under such conditions how to enable the dynamic execution model to progress on a sub-task level? \textit{contribution---3} To avoid non-termination under adversarial
energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
Partial execution avoids non-termination by committing the intermediate state of a long-running
task that has repeatedly failed and restarted. Partial execution violates
task atomicity, but preserves forward progress; if a programmer knows that
task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
\end{itemize}
%
Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} provides an overview of \sys, while
Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism. Dynamic group privatization mechanism is provided in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's methodology and evaluation, respectively. Section~\ref{sec:related_work} puts \sys in the
context of related work and Section~\ref{sec:conclusions} concludes and discusses
future work.
